### 2014-03-14 src/methodology.Rnw
methodology alex

\section{Methodology}
\label{sec:methodology}
Our selection of programs was driven by a few overriding concerns. Our primary
requirement was that, our results had to be as widely applicable as possible
for real-world programs. Secondly, we strived for a statistically significant
result, reducing the amount of variables present in the experiments. For this
reason we chose a random sample of Java projects from Github\cite{github}
that uses the popular maven\cite{maven} build system, that has been used
elsewhere\cite{gopinath2014code}. We utilize two flavors of
of test-suites. The original test-suites present in a limited number of
projects, and automatically generated test-suites using randoop.  Due
limited resources, we had to further reduce the programs and test-suites
to those that could finish the mutation run in 10 minutes. From this set
of test-suites, we sampled 110 suites each for both original and randoop
generated test-suits. This gave us 81 unique projects for both original
and randoop, and remaining projects having both flavors of test suites.

We ran mutation analysis on this set of projects using
PIT\cite{pitest}. However, since the operators provided by PIT are limited,
we extended PIT to provide newer operators that are similar to operators
provided by other mutation systems. The set of operators that we used
are provided in Table~\ref{tbl:pitop}. A distribution of these operators
in our projects are provided in Figure~\ref{fig:op}, which shows both the
distribution of complete mutants and the relative detection rates of each
mutation operators. In the figure, there are two box-plots corresponding
to each mutation operator on the Y-axis. The black box represents the
relative frequency of the particular mutation operator in the total mutants
\textit{generated}. For example for projects in the randoop set, the
operator \pitnvmc contributed about 34\% of the total number of mutants
produced while \pitrv contributed about 7\%. Similarly, the blue shaded
box-plot represents the relative frequency of particular mutation operator
in the total mutants \textit{detected}. For example, \pitnvmc contribution
in the detected mutants was 23\% while that of \pitrv was 15\%.

For a detailed description of each mutants, please see~\cite{pitmut}.
PIT was also modified to provide random selection of mutants according
to various criteria. Finally, to remove random noise, each criteria
described has been run four times, and the results are averaged to produce
the final result.

@@

One of the criteria we sampled was \textit{statement deletion}.
Since we use PIT, which is a bytecode based mutation engine, implementing direct equivalent of
source code statement deletion is very hard. We were able to approximate statement deletion
by combining different bytecode operators, which is explained in Subsection~\ref{subsec:stmtdel}.

@@

We used several different sampling criteria, some of which has been suggested
in the literature before, and some which are variants of previously suggested
criteria, and some that have not been used before.  For each sampling
criteria, we sampled mutants on a decreasing power scale sampling
\textit{1, 1/2, 1/4, 1/8, 1/16, 1/64} of the total mutants.

\subsubsection{x\% selection}
The simplest one consisted of using x\% selection as suggested by
Budd~\cite{budd1980mutation}. In this criteria, we choose a specific
fraction of the complete set of mutants. This criteria also serves as a
control for the verifying the effectiveness of other criteria.

\subsubsection{Sampling over program elements (and variations)}
Following the suggestion of Zhang et al.~\cite{zhang2013ase}, we
extended x\% selection criteria to sample from within different program
elements.  We sampled in the increasing order of scope,
--- \textit{line}, \textit{method} and \textit{class} (\textit{project}
scope was analyzed as part of x\% selection).

@@

Next, we modified this criteria slightly, and instead of choosing a mutant
with probability of decimal part, applied \textit{round(sample)} to round
it to the nearest whole number. We used this as the number of samples to
be chosen from the population. Our aim was to remove elements that was
too simple. That is, if the number of mutants dropped below a threshold
determined by the mutant reduction ratio, that program element did not
contribute any thing to the final result.

Next, we explored in the other direction, by forcing the program elements
to return at least one element by using the \textit{ceil(sample)}. That is,
irrespective of the fraction being sampled, the simplest elements always
contributed at least one mutant to the total sample.

\subsubsection{Lines per element and variants}
Our previous research~\cite{gopinath2014code} found that statement coverage
was highly correlated with mutation coverage for a project. This immediately
suggests that perhaps choosing one mutant per line may be sufficient to achieve
a close approximation of the final mutation score. Further, statement deletion
has been researched previously~\cite{untch2009onreduced,deng2013empirical} and
has been found to reduce the number of mutants well, without an appreciable
decrease in accuracy. This also provides a nice comparison with similar
number of mutants between operator selection and random sampling of mutants.

We next extended it to sample from each method instead, but using the number
of lines in the method as the number of mutants to be sampled. This provided
a test of hypothesis of how important the coverage was. This was extended
to class and project scope also.

For the first variant, we applied the x\% sampling to the result of the first
sampling based on line counts. That is, in the case of methods, we first
selected \textit{linecount} mutants each from each method, and applied the x\%
selection to the result of this selection. For the second variant, we did
the sampling in one go, where we chose \textit{linecount/x} number of mutants from
each program element.

@@

For selective methods, we tried mutation operators suggested by Wong et al.~\cite{wong1995reducing}, Offutt et al.~\cite{offutt1996anexperimental,deng2013empirical}, % TODO Barbosa
 and, Namin et al~\cite{siami2008sufficient}. Since Javalanche~\cite{javalanche} utilized operator selection mechanisms, we also compared the Javalanche operators for operator selection. Note that all of these techniques except 
Javalanche have targeted C programs, thus some of these operators
are sensible in C but in Java programs. For example C compilers tolerate
removing a return statement but, in Java, it causes a compilation error.
Moreover, there were some operators that are not supported by PIT framework
like ABS (absolute mutation operators), however, we tried to realize as many
as possible operators using PIT's bytecode level mutation. We also extended 
the PIT framework to support the  statement deletion operator (SDL).

>>>>

\section{Methodology} \label{sec:methodology} Our selection of
programs was driven by a few overriding concerns. Our primary
requirement was that our results had to be as widely applicable as
possible for real-world programs. Secondly, we strived for a
statistically significant result, reducing the number of variables
present in the experiments. For this reason we chose a random sample
of Java projects from Github\cite{github} that use the popular
maven\cite{maven} build system, following the methodology used by some
of the authors in an ICSE 2014 paper~\cite{gopinath2014code}. As in
that work, we utilize two kinds of of test suites: the original test
suites present in a limited number of projects, and automatically
generated test suites from Randoop~\cite{randoop}.  Due to limited resources, we had to
further reduce the programs and test suites to those that could finish
mutation analysis in 10 minutes. From this set of test suites, we
sampled 110 suites each for both original and randoop generated
test suites. This gave us 81 unique projects for both original and
Randoop suites, and remaining projects having both flavors of test suites. {\bf ADG I have no idea what that last sentence means, really.}

We ran mutation analysis on this set of projects using
PIT\cite{pitest}. However, since the operators provided by PIT are limited,
we extended PIT to provide newer operators that are similar to operators
provided by other mutation systems. The set of operators that we used
is provided in Table~\ref{tbl:pitop}. Figure~\ref{fig:op} which shows both the
distribution of mutants and the relative detection rates of each
mutation operator for our projects. In the figure, there are two box-plots corresponding
to each mutation operator on the Y-axis. The black box represents the
relative frequency of the particular mutation operator in the total mutants
\textit{generated}. For example for projects in the Randoop set, the
operator \pitnvmc contributed about 34\% of the total number of mutants
produced while \pitrv contributed about 7\%. Similarly, the blue shaded
box-plot represents the relative frequency of particular mutation operator
in the total mutants \textit{detected}. For example, \pitnvmc contribution
in the detected mutants was 23\% while that of \pitrv was 15\%.

For a detailed description of each mutant type, please see~\cite{pitmut}.
PIT was also modified to provide random selection of mutants according
to various criteria. Finally, to remove random noise, each criteria
described was run four times, and the results averaged to produce
the final result.

@@

One of the criteria we sampled was \textit{statement deletion}.
Since we use PIT, which is a bytecode based mutation engine, implementing direct equivalent of
source code statement deletion is very hard. We were able to approximate statement deletion
by combining different bytecode operators, as explained in Subsection~\ref{subsec:stmtdel}.

@@

We used several different sampling criteria, some of which has been suggested
in the literature before, some of which are variants of previously suggested
criteria, and some of which have not been used before.  For each sampling
criteria, we sampled mutants on a decreasing power scale, sampling
\textit{1, 1/2, 1/4, 1/8, 1/16, 1/64} of the total mutants.

\subsubsection{x\% selection}
The simplest sampling approach consisted of using x\% selection as suggested by
Budd~\cite{budd1980mutation}. In this criteria, we choose a specific
fraction of the complete set of mutants. This criteria also serves as a
control for verifying the effectiveness of other criteria.

\subsubsection{Sampling over program elements (and variations)}
Following the suggestion of Zhang et al.~\cite{zhang2013ase}, we
extended x\% selection criteria to sample from within different program
elements.  We sampled in increasing order of scope,
--- \textit{line}, \textit{method} and \textit{class} (\textit{project}
scope is just x\% selection).

@@

Next, we modified this criteria slightly, and instead of choosing a
mutant with probability equal to the decimal part, we applied
\textit{round(sample)} to round it to the nearest whole number. We
used this as the number of samples to be chosen from the
population. Our aim was to \emph{remove} elements that were too simple. That
is, if the number of mutants dropped below a threshold determined by
the mutant reduction ratio, that program element did not contribute
any mutants to the final result.  Conceptually the idea is that mutants of more complex code have more discriminatory power (and may correspond better to real faults though we do not evaluate this concept).

Next, we explored in the other direction, by forcing the program elements
to return at least one element by using \textit{ceil(sample)}. That is,
irrespective of the fraction being sampled, the simplest elements always
contributed at least one mutant to the total sample.

\subsubsection{Lines per element and variants}
Our previous research~\cite{gopinath2014code} found that statement coverage
was highly correlated with mutation coverage for a project. This immediately
suggests that perhaps choosing one mutant per line may be sufficient to achieve
a close approximation of the final mutation score. Further, statement deletion
has been researched previously~\cite{untch2009onreduced,deng2013empirical} and
has been found to reduce the number of mutants well, without an appreciable
decrease in accuracy. This also provides a nice comparison with similar
numbers of mutants between operator selection and random sampling of mutants.

We next extended this approach to sample from each method instead, but
used the number of lines in the method as the number of mutants to be
sampled from it. This provided a test of the hypothesis of how
important a mutation samplings relationship to simple code coverage
was. This was extended to class and project scope also.

For the first variant, we applied the x\% sampling to the result of the first
sampling based on line counts. That is, in the case of methods, we first
selected \textit{linecount} mutants each from each method, and applied the x\%
selection to the result of this selection. For the second variant, we did
the sampling in one go, where we chose \textit{linecount/x} number of mutants from
each program element. {\bf ADG:  Could maybe make this clearer?}

@@

For selective methods, we tried mutation operators suggested by Wong et al.~\cite{wong1995reducing}, Offutt et al.~\cite{offutt1996anexperimental,deng2013empirical}, % TODO Barbosa
 and, Namin et al~\cite{siami2008sufficient}. Since Javalanche~\cite{javalanche} utilized operator selection mechanisms, we also compared the Javalanche operators for operator selection. Note that all of these techniques except 
Javalanche have targeted C programs, thus some of these operators
are sensible in C but not in Java programs. For example C compilers tolerate
removing a return statement --- in Java this causes a compilation error.
Moreover, there were some operators that are not supported by the PIT framework
like ABS (absolute mutation operators). We tried to realize as many
as possible of the operators using PIT's bytecode level mutation. We also extended 
the PIT framework to support the  statement deletion operator (SDL).
